---
title: "2015 311 Data"
author: "Ryan Elmore & Kellie Keeling"
date: "June 20, 2016"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: default
    highlight: tango
---
```{r packages, include=FALSE}
library(knitr)
```

```{r setup, echo=FALSE}
opts_chunk$set(cache=TRUE,
               echo=FALSE,
               warning = FALSE,
               message = FALSE,
               comment = NA)
```

```{r load-data}
df <- readRDS(gzcon(url("https://github.com/rtelmore/rmd_examples/raw/master/data/311-df.rds")))

```

##Exploration of the 2015 Denver 311 Dataset
This report explores a 10% sample of the 2015 311 data by using visualizations.

### Line Charts
The first chart we have shows a line chart showing the number of incidents by month.  We have defined several options such as changing the default axis showing months 1 - 12 to letters, changing the y axis label, and adding a title. The second chart shows the Number of incidents by hour of day where 2 vertical lines have been added with text that show the 

```{r line-charts}

#RESOURCE: http://www.statmethods.net/index.html
#Line Charts
par(mar=c(5.1, 4.1, 4.1, 2.1)) #default #bottom, left, top and right margins
plot(table(df$Month),
     main="The Number of Incidents for 2015",
     ylab = "Number of Incidents",
     type="l",
     xaxt="n")
axis(side=1, 
     at=c(1,2,3,4,5,6,7,8,9,10,11,12),
     labels=c("J","F","M","A","M","J","J","A","S","O","N","D"))

plot(table(df$Hour),
     main="Hour of Day",
     ylab = "Number of Incidents",
     type="l")
abline(v=8)
abline(v=17)
text(x=8, 
     y=4700, 
     labels="8:00am", 
     cex = .8)
text(x=18, 
     y=4700, 
     labels="5:00pm", 
     cex = .8)
```

### Column Chart
The next charts shows an example of a column chart.  Here we can see that most of the calls were not resolved on the first call and the title displays the total number of Y and N values from the data so it will update with a new data set.

```{r column-chart}
#Column Chart
barplot(sort(table(df$First.Call.Resolution)),
        main=paste("First Call Resolution Y:", 
                   table(df$First.Call.Resolution)[2],
                   "N:",
                   table(df$First.Call.Resolution)[3]))
```


```{r}

```

### Bar Plot
Next we see the example of 2 bar plots.  Bar plots are preferred over column charts when you have long names for the bars.

```{r bar-plot}

#Bar Plots
par(mar=c(5.1,12.1,4.1, 2.1)) #make left margin larger
barplot(sort(table(df$Case.Status)),
        horiz = TRUE,
        cex.names=.75, 
        las=1,
        main="Case Status" ) #las=1 horiz labels
par(mar=c(5.1, 4.1, 4.1, 2.1)) #default #bottom, left, top and right margins

#This plot doesn't need the extra left margin
barplot(sort(table(df$Council.District)),
        horiz = TRUE, 
        cex.names=.75, 
        las=1,
        main="Council District")

```

### Histogram
 
The final example in this file shows a Histogram. It is fairly plan with just the default output.
```{r histogram}
#Histogram
x <- as.numeric(df$DaysToClose[df$DaysToClose<60])
hist(x)
```

## Exercises:  

1. Try changing the Hour of Day line chart to show vertical lines and 7:00 am and 6:00pm.  
2. Create bar plots for Agency, Type, or Police District. 
3. Try making the histogram graph more complete. 
4. Some options to try:  
* xlab (xlabels) 
* main (title) 
* sub (bottom title) - try putting mean in subtitle using code: paste("mean:",mean(x,na.rm=TRUE))


## Appendix: The Code
```{r appendix, echo=TRUE, eval=FALSE, ref.label=all_labels()}
```
